---
title: "Methylation"
author: "Xiangning Xue"
date: "5/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(GEOquery)
library(Biobase)
library(dplyr)
```

## Prepare data
```{r meta_data}
GEO.ID = "GSE103249"
gset <- getGEO(GEO.ID,  GSEMatrix =TRUE, getGPL=TRUE)
data.pdata = pData(gset[[1]]) #the meta data
# summary of the data: all male tissues from cerebral cortex (18 WT + 18 Snord116+/-)
ZT.time = gsub(".*ZT([0-9]+)_rep[0-9]", "\\1", data.pdata$title) %>% as.numeric()
group.label = gsub("(.*) ZT.*", "\\1", data.pdata$title)
write.csv(data.pdata[, "supplementary_file_1", drop = FALSE], "RawDataFTP.csv") #delete header manually
meta.df = data.frame(ZT = ZT.time, genotype = group.label)
rownames(meta.df) = rownames(data.pdata)
write.csv(meta.df, "metadata.csv")
```

```{r exprs_data exploration, eval=FALSE}
######## The following does not work
# #!/bin/sh
# while IFS=, read -r sample ftp
# do
#     ftp --recursive --no-parent -nd -O ./raw/$sample.zip $ftp
# done < RawDataFTP.csv

######## The following works
# wget 'ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE103nnn/GSE103249/suppl/GSE103249_RAW.tar'
# cd raw
# tar –xvzf raw/GSE103249_RAW.tar
hdir = "/home/xix66/circadian/ThePipeline/CircadianPipeline_private/RealData/methylation"
bed <- as.data.frame(read.table(paste0(hdir, "/raw/GSM2758987_JLRC002B.bed.gz"),header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
colnames(bed) = c("chrom", "chromStart", "chromEnd", "V4", "V5", "strand", "V7", "V8", "V9")
#some explorations
dim(bed) #15223218        9
table(bed$strand) #all "+", so just one strand
quantile(bed$V5, na.rm =TRUE) # all 0
quantile(bed$V7, na.rm =TRUE) # all 0
quantile(bed$V8, na.rm =TRUE) # all 0
# V5, V7, V8 are placeholders. 
#V4; percent methylation and number of reads contributing to the permeth call
bed$pm = gsub("(.*)-.*", "\\1",  bed$V4)
bed$nReads = gsub(".*-(.*)", "\\1",  bed$V4)
quantile(as.numeric(bed$pm), na.rm =TRUE) 
#  0%  25%  50%  75% 100% 
#0.00 0.67 1.00 1.00 1.00 
quantile(as.numeric(bed$nReads), na.rm =TRUE) 
#  0%  25%  50%  75% 100% 
#   1    1    3    4  147
bed2 <- as.data.frame(read.table(paste0(hdir, "/raw/GSM2758975_JLRC002A.bed.gz"),header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
dim(bed)#15223218       10
dim(bed2)#15573035        9
bed2$pm = gsub("(.*)-.*", "\\1",  bed2$V4)
bed2$nReads = gsub(".*-(.*)", "\\1",  bed2$V4)
quantile(as.numeric(bed2$pm), na.rm =TRUE) 
#  0%  25%  50%  75% 100% 
#0.00 0.67 1.00 1.00 1.00 
quantile(as.numeric(bed2$nReads), na.rm =TRUE) 
#  0%  25%  50%  75% 100% 
#   1    2    3    5  150
sum(as.numeric(bed2$nReads)>10) 
sum(as.numeric(bed$nReads)>10)
head(bed2[bed2$nReads==1, ], n = 10)
head(bed[bed$nReads==1, ], n = 10)
head(bed2[bed2$nReads==2, ], n = 10)
head(bed[bed$nReads==2, ], n = 10)
```


```{r exprs_data process the data, eval=FALSE}
ToBeta = function(aBedSample){
  V4 = aBedSample$V4
  chromInfo = aBedSample[, 1:3]
  pm = gsub("(.*)-.*", "\\1",  V4)
  nReads = gsub(".*-(.*)", "\\1",  V4) 
  return(list(pm = cbind.data.frame(chromInfo, pm = pm),
              nReads = cbind.data.frame(chromInfo, nReads = nReads)))
}

samples = list.files(paste0(hdir, "/raw"))
samples = samples[samples != "GSE103249_RAW.tar"]
ALL.BED.INFO = parallel::mclapply(samples, function(a){
  a.bed <- as.data.frame(read.table(paste0(hdir, "/raw/", a),header = FALSE, sep="\t",stringsAsFactors=FALSE, quote=""))
  res = ToBeta(a.bed)
  return(res)
}, mc.cores = 18)
names(ALL.BED.INFO) = gsub("(GSM[0-9]+)_.*", "\\1", samples)
saveRDS(ALL.BED.INFO, paste0(hdir, "/raw/AllBedInfo.rds"))
# ALL.pm = Reduce(function(x, y) merge(x, y, by=c("V1", "V2", "V3")), lapply(ALL.BED.INFO, `[[`, 1))
# ALL.nReads = Reduce(function(x, y) merge(x, y, by=c("V1", "V2", "V3")), lapply(ALL.BED.INFO, `[[`, 2))
# dim(All.pm)


```

```{r buid_bsseq_obj_with_github_code}
library(bsseq) 
#original code see: https://github.com/hansenlab/bsseq/blob/master/inst/scripts/get_BS.chr22.R
read.mydat <- function(a.file, SampleName) {
  #my date only has one strand
    BS <- BSseq(
        pos = a.file[[1]]$V2,
        chr = a.file[[1]]$V1,
        M = as.matrix(as.integer(as.numeric(as.character(a.file[[1]]$pm))*as.numeric(as.character(a.file[[2]]$nReads))), ncol = 1),
        Cov = as.matrix(as.integer(as.numeric(as.character(a.file[[2]]$nReads))), ncol = 1),
        sampleNames = SampleName)
    BS
}
BS.list = lapply(1:length(ALL.BED.INFO), function(a){
  a.BS <- read.mydat(ALL.BED.INFO[[a]], names(ALL.BED.INFO)[a])
  sampleNames(a.BS) <- names(ALL.BED.INFO)[a]
  return(a.BS)
  })

BS.obj <- combineList(BS.list)
BS.obj <- orderBSseq(BS.obj, seqOrder = c(paste0("chr", seq(1:19)), "chrM", "chrX", "chrY"))
head(getCoverage(BS.obj, type = "M"), n = 4)[, 1:3]
head(getCoverage(BS.obj, type = "Cov"), n = 4)[, 1:3]
BS.temp = BS.list[[1]]
BS.temp <- orderBSseq(BS.temp, seqOrder = c(paste0("chr", seq(1:19)), "chrM", "chrX", "chrY"))
head(getCoverage(BS.temp, type = "Cov"), n = 4)
granges(BS.obj)[1:4, ]
granges(BS.temp)[1:4, ]
ALL.BED.INFO[[1]][[1]][ALL.BED.INFO[[1]][[1]]$V1=="chr1"&ALL.BED.INFO[[1]][[1]]$V2==3000827, ]
ALL.BED.INFO[[1]][[2]][ALL.BED.INFO[[1]][[1]]$V1=="chr1"&ALL.BED.INFO[[1]][[1]]$V2==3000827, ]

meta = read.csv(paste0(hdir, "/metadata.csv"), row.names = 1)
all(names(ALL.BED.INFO)%in%rownames(meta)) #TRUE
all(rownames(meta)%in%names(ALL.BED.INFO)) #TRUE
meta = meta[names(ALL.BED.INFO), ]
pData(BS.obj)$ZT <- meta$ZT
pData(BS.obj)$genotype <- meta$genotype
validObject(BS.obj)
pData(BS.obj)
saveRDS(BS.obj, paste0(hdir, "/BSobj.rds"))
# rm(bed, bed2, ALL.BED.INFO, BS.list)
```

From the raw data, we get a methylation percentage table of dimension (), and a read number table of the same dimension. We first filter the CpG loci to contain only those that have reads in all samples, which leaves us () loci. 

Two statistics used ($N_{i, \text{methy}}$ is the number of methylated reads in sample $i$, and $N_{i, \text{unmethy}}$ is the number of unmethylated reads in sample $i$): 
  
  * Beta value: $B_i = \frac{N_{i, \text{methy}}}{N_{i, \text{unmethy}}+\alpha}$,
  * M value: $M_i = \log_2{\frac{N_{i, \text{methy}}+\alpha}{N_{i, \text{unmethy}}+\alpha}}$. 
Here $\alpha$ is a offset constant. 
  
Some loci has low reads in some samples, thus the methylation percentage evaluated might not be accurate. Specially, loci with #reads = 1 will have beta value = 0 or 1 for all samples. 

### Reload the data
```{r}
hdir = "/home/xix66/circadian/ThePipeline/CircadianPipeline_private/RealData/methylation"
BS.obj = readRDS(paste0(hdir, "/BSobj.rds"))
```


## Circadian analysis

### Fit by CpG site.
```{r, eval = FALSE}
# if we fit model locus by locus, problem is that the curve is not smooth enough. Also it will be harder to explain. 
# Also, the result shows weak signal, better do something to strengthen the signal. 

#filter out loci with too many zero reads
filterByNReads = function(groups, nReads){
  nReads.list = sep(t(nReads), groups)
  reads.sum
}
keep = apply(getCoverage(BS.obj, type = "Cov"), 1, function(x){mean(x>0)}) #the kept loci must have at least one read in each sample
#6970239 out of 20235445
keep2 = apply(getCoverage(BS.obj, type = "Cov"), 1, function(x){mean(x>1)}) #the kept loci must have at least two read in each sample
#2791512
keep5 = apply(getCoverage(BS.obj, type = "Cov"), 1, function(x){mean(x>=5)}) #the kept loci must have at least two read in each sample
#97133
#Try fit with threshold 2, but actually should use 5. 
BS.obj.sub = BS.obj[keep2 == 1, ]
meth.sub = getMeth(BS.obj.sub, type = "raw")
M.sub = log2((meth.sub+1e-5)/(1-meth.sub+1e-5))

x1 = list(data = data.frame(M.sub[, pData(BS.obj.sub)$genotype == "WT"]), 
          time = pData(BS.obj.sub)[pData(BS.obj.sub)$genotype == "WT", "ZT"], 
          gname = seq_len(nrow(M.sub)))
x2 = list(data = data.frame(M.sub[, pData(BS.obj.sub)$genotype != "WT"]), 
          time = pData(BS.obj.sub)[pData(BS.obj.sub)$genotype != "WT", "ZT"], 
          gname = seq_len(nrow(M.sub)))
DCP_rhythm = DiffCircaPipeline::DCP_Rhythmicity(x1, x2, parallel.ncores = 20)
time1 = Sys.time()
DCP_dparam = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "A&phase")
time2 = Sys.time()
DCP_dR2 = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm)
time3 = Sys.time()
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_rhythm$rhythm.joint %>% 
                                     filter(TOJR=="both")%>%
                                     mutate(meanP = 1/2*sum(c(1-log(pG1), 1-log(pG2)))) %>%
                                     arrange(meanP) %>%
                                     pull(gname) %>% head(20), 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Scatter_both", file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Scatter_both_1", file.height = 8, file.width = 16)
time4 = Sys.time()
DiffCircaPipeline::DCP_PlotPeakDiff(DCP_rhythm, NULL, DCP_dparam,
                                    color.cut = list(param = "post.hoc.peak.By.p", fun = ">", val = 0, color.sig = "#b33515", color.none = "dark grey"),
                                    Info1 = "WT", Info2 = "Snord116",
                                    filename = "PhaseDiff", file.height = 8, file.width = 8, time.start = 0)
time5 = Sys.time()
DiffCircaPipeline::DCP_PlotPeakRadar(DCP_rhythm, NULL, TRUE,
                                     Info1 = "WT", Info2 = "Snord116",
                                     filename = "PhaseRadar", file.height = 8, file.width = 8, time.start = 0)
time6 = Sys.time()
```


```{r methylsig, eval = FALSE}
# test smooth by window. But currently I want to smooth by gene because it is easier to explain. 
# BiocManager::install("methylSig")
# BiocManager::install("DSS")
# install.packages("/home/xix66/circadian/ThePipeline/CircadianPipeline_private/methylSig_1.8.0.tar", repos = NULL)

#filter out low reads. (will not filter by SNP because I do not know the genotype)
BS.obj.msfilter = methylSig::filter_loci_by_coverage(BS.obj, min_count = 5, max_count = 500)
#check dimension: both still have 20235445 methylation loci, maybe just the low ones are set to 0 but not deleted. 

#test tiling by window (use a large window)
windowed_bs = methylSig::tile_by_windows(bs = BS.obj.msfilter, win_size = 100000)
windowed_meth = getMeth(windowed_bs, type = "raw")
#how many does not contain NA?
windowed_meth.na = parallel::mclapply(1:nrow(windowed_meth), function(a){
  a.row = as.numeric(windowed_meth[a, ])
  mean(is.na(a.row))
  }, mc.cores = 20)
sum(windowed_meth.na==0)#25324 still plenty 


```

### Fit by gene
```{r gene_windows}
#Mus musculus
# BiocManager::install("methylKit")
# BiocManager::install("genomation")
# I downloaded a map here: http://genome.ucsc.edu/cgi-bin/hgTables
gene.obj0 = read.csv("/home/xix66/circadian/ThePipeline/CircadianPipeline_private/RealData/methylation/Galaxy3-[UCSC_Main_on_Mouse__wgEncodeGencodeCompVM11_(genome)].csv")
gene.obj0 = gene.obj0[, c(2:6, 13)]#should keep txStart and txEnd
gene.obj0 = gene.obj0[gene.obj0$strand=="+", ]
gene.obj0 = gene.obj0[gene.obj0$chrom %in% c(paste0("chr", seq(1:19)), "chrM", "chrX", "chrY"), ]
gene.obj0 = gene.obj0[, -3]
#keep unique genes and keep the one with largest range
time1 = Sys.time()
gene.obj0.by.gene = split(gene.obj0, list(as.character(gene.obj0$name2))) %>% #19473 genes
  lapply(function(a.gene.tab){
    if(nrow(a.gene.tab)<=1){
      keep = a.gene.tab
    }else{
      width = a.gene.tab$txEnd-a.gene.tab$txStart
      keep = a.gene.tab[which.max(width),]
    }
    return(keep)
  })
time2 = Sys.time()
gene.obj0.unique.gene = do.call(rbind.data.frame, gene.obj0.by.gene)
dim(gene.obj0.unique.gene) #19473
range = gene.obj0.unique.gene$txEnd-gene.obj0.unique.gene$txStart
quantile(range)
     # 0%     25%     50%     75%    100% 
     # 11    1067    6226   25717 4434882 

gene.obj=GenomicRanges::makeGRangesFromDataFrame(gene.obj0.unique.gene, keep.extra.columns = TRUE, ignore.strand=TRUE,
                                                 seqnames.field = "chrom", start.field = "txStart", end.field = "txEnd", 
                                                 strand.field = "strand")#19473 ranges

BS.obj.msfilter = methylSig::filter_loci_by_coverage(BS.obj, min_count = 5, max_count = 500)
# #check what happens after filtering by coverage. -> Basically a clear cut off, if less than 5 then set as 0. 
# BS.obj.Cov = bsseq::getCoverage(BS.obj, type = "Cov")
# BS.obj.msfilter.Cov = bsseq::getCoverage(BS.obj.msfilter, type = "Cov")
# quantile(BS.obj.Cov)
#   # 0%  25%  50%  75% 100% 
#   #  0    1    3    5  177
# quantile(BS.obj.msfilter.Cov)
#  #  0%  25%  50%  75% 100%  
#  #   0    0    0    5  177 
# mean(BS.obj.Cov!=0) #0.3072646
# mean(BS.obj.msfilter.Cov!=0) #0.8590005
region_bs = methylSig::tile_by_regions(bs = BS.obj.msfilter, gr = gene.obj)#19473 
meth.region = bsseq::getMeth(region_bs, type = "raw")
M.region = log2((meth.region+1e-5)/(1-meth.region+1e-5))
p.na = lapply(1:nrow(M.region), function(a){
  a.row = as.numeric(M.region[a, ])
  mean(is.na(a.row))
  })
p.na = unlist(p.na)
sum(p.na==0) #19234
sum(p.na==0) #19234
M.region.keep = M.region[p.na==0, ] #12815
meth.region.keep= bsseq::getCoverage(region_bs, type = "M")[p.na==0, ]
cov.region.keep= bsseq::getCoverage(region_bs, type = "Cov")[p.na==0, ]
gname.keep = as.data.frame(GenomicRanges::granges(region_bs))[p.na==0, ]

x1 = list(data = data.frame(M.region.keep[, bsseq::pData(region_bs)$genotype == "WT"]), 
          time = bsseq::pData(region_bs)[bsseq::pData(region_bs)$genotype == "WT", "ZT"], 
          gname = gname.keep$name2)
x2 = list(data = data.frame(M.region.keep[, bsseq::pData(region_bs)$genotype != "WT"]), 
          time = bsseq::pData(region_bs)[bsseq::pData(region_bs)$genotype != "WT", "ZT"], 
          gname = gname.keep$name2)
DCP_rhythm = DiffCircaPipeline::DCP_Rhythmicity(x1, x2)
table(DCP_rhythm$rhythm.joint$TOJR)
table(DCP_rhythm$rhythm.joint$TOJR.FDR)
DCP_dparam = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "A&phase")
DCP_dparam2 = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "phase")
DCP_dR2 = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm)
save(x1, x2, DCP_rhythm, DCP_dparam, DCP_dparam2, DCP_dR2, file = paste0(hdir, "/DCP_output.rData"))
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_rhythm$rhythm.joint %>% 
                                     filter(TOJR=="both")%>%
                                     mutate(tG1 = 1-log(pG1), tG2 = 1-log(pG2)) %>%
                                     mutate(meanP = mean(c(tG1, tG2))) %>%
                                     arrange(meanP) %>%
                                     pull(gname) %>% head(20), 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = paste0(hdir, "/ByGene_Scatter_both"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = paste0(hdir, "/ByGene_Scatter_1.pdf"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dparam %>%
                                     filter(post.hoc.peak.By.p) %>% 
                                     pull(gname),
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = paste0(hdir, "/ByGene_Scatter_dPhase"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dR2 %>%
                                     arrange(p.R2) %>% 
                                     pull(gname) %>% 
                                     `[`(1:20),
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = paste0(hdir, "/ByGene_Scatter_dR2"), file.height = 8, file.width = 16)

DiffCircaPipeline::DCP_PlotPeakDiff(DCP_rhythm, NULL, NULL,
                                    color.cut = list(param = "post.hoc.peak.By.p", fun = "<", val = 1, color.sig = "#b33515",
                                                     color.none = "dark grey"),
                                    Info1 = "WT", Info2 = "Snord116",
                                    filename = paste0(hdir, "/ByGene_PhaseDiff.pdf"), file.height = 8, file.width = 8, time.start = 0)
DiffCircaPipeline::DCP_PlotPeakRadar(DCP_rhythm, NULL, TRUE,
                                     Info1 = "WT", Info2 = "Snord116",
                                     filename = paste0(hdir, "/ByGene_PhaseRadar.pdf"), file.height = 8, file.width = 8, time.start = 0)

load("DCP_output.rData")
load(paste0(hdir, "/DCP_output.rData"))
```

#### Validating the result with permutation test

Since the result shows very weak signal, it is possible that the model does not fit the data. We validate the performance with permutation and the p-values from the two methods are very concordant. 
```{r}
library('minpack.lm')
source(paste0(hdir, '/code/fitSinCurve.R'))

dir.create(paste0(hdir, "/FitByGene_PermRes"), recursive = TRUE)
B<-1000
set.seed(12345)
result = parallel::mclapply(1:B, function(b){
  print(b)
  x1.perm = x1; x1.perm$time = sample(x1.perm$time)
  x2.perm = x2; x2.perm$time = sample(x2.perm$time)
  DCP_rhythm.x1.perm = DiffCircaPipeline::DCP_Rhythmicity(x1.perm)
  DCP_rhythm.x2.perm = DiffCircaPipeline::DCP_Rhythmicity(x2.perm)
  a.res = list(x1.perm.res = DCP_rhythm.x1.perm$rhythm, 
              x2.perm.res = DCP_rhythm.x2.perm$rhythm)
  saveRDS(a.res, paste0(hdir, "/FitByGene_PermRes", "/", "perm", b, ".rds"))
  return(a.res)
}, mc.cores = 20)

result = parallel::mclapply(1:B, function(b){
  readRDS(paste0(hdir, "/FitByGene_PermRes", "/", "perm", b, ".rds"))
}, mc.cores = 20)

n <- nrow(x1$data)
x1_null_R2 <- matrix(0,n,B)
x2_null_R2 <- matrix(0,n,B)

for(b in 1:B){
  print(b)
  x1_null_R2[, b] <- result[[b]][[1]]$R2
  x2_null_R2[, b] <- result[[b]][[2]]$R2
}

x1_R2_pool <- c(DCP_rhythm$x1$rhythm$R2, x1_null_R2)
x1_R2Rank <- 1 - (rank(x1_R2_pool)[1:length(DCP_rhythm$x1$rhythm$R2)] - 0.5)/length(x1_R2_pool)
x2_R2_pool <- c(DCP_rhythm$x2$rhythm$R2, x2_null_R2)
x2_R2Rank <- 1 - (rank(x2_R2_pool)[1:length(DCP_rhythm$x2$rhythm$R2)] - 0.5)/length(x2_R2_pool)

pdf(paste0(hdir, "/FitByGene_CompareWithPermRes.pdf"))
par(mfrow=c(1, 2))
plot(-log10(DCP_rhythm$x1$rhythm$pvalue), -log10(x1_R2Rank), xlab = "DCP pvalues", ylab = "permutation pvalues", main = "WT")
abline(a = 0, b = 1, col = "red", lty = 2)
plot(-log10(DCP_rhythm$x2$rhythm$pvalue), -log10(x2_R2Rank), xlab = "DCP pvalues", ylab = "permutation pvalues", main = "Snord116")
abline(a = 0, b = 1, col = "red", lty = 2)
dev.off()
```

#### Enrichment analysis
```{r}
#speed 4
hdir = "/home/xix66/circadian/ThePipeline/CircadianPipeline_private/RealData/methylation"
hdir = "/Users/xiangningxue/Library/CloudStorage/OneDrive-UniversityofPittsburgh/Research/circadian/ThePipeline/R_package_github/DiffCircaPipeline/MethylationData"
background.genes = x1$gname
length.max = length(background.genes)
fill.vec = function(a.vec, length.max=nrow(res)){
  c(a.vec, rep("", length.max-length(a.vec)))
}
df = data.frame(RhyI = fill.vec(as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR %in% c("rhyI")], length.max),
                RhyII = fill.vec(as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR %in% c("rhyII")], length.max),
                RhyBoth = fill.vec(as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR=="both"], length.max),
                Background = background.genes)
colnames(df) = c("WT only", "Snord116 only", "overlaps", "_BACKGROUND")
dir.create(paste0(hdir, "/MetaScape_Input"))
write.csv(df, paste0(hdir, "/MetaScape_Input/DCP_byGene_TOJR_", "WT", "_", "Snord116", ".csv"), row.names = FALSE)

df2 = data.frame(WT = fill.vec(as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR %in% c("rhyI", "both")], length.max),
                Snord116 = fill.vec(as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR %in% c("rhyII", "both")], length.max),
                Background = background.genes)
colnames(df2) = c("WT", "Snord116", "_BACKGROUND")
write.csv(df2, paste0(hdir, "/MetaScape_Input/DCP_byGene_TOJR2_", "WT", "_", "Snord116", ".csv"), row.names = FALSE)

#differential phase genes 
df3 = data.frame(R2up = fill.vec(as.character(DCP_dR2$gname)[DCP_dR2$p.R2<0.05&DCP_dR2$delta.R2>0], length.max),
                 R2down = fill.vec(as.character(DCP_dR2$gname)[DCP_dR2$p.R2<0.05&DCP_dR2$delta.R2<0], length.max),
                 Background = background.genes)
colnames(df3) = c("R2up", "R2down", "_BACKGROUND")
write.csv(df3, paste0(hdir, "/MetaScape_Input/DCP_byGene_dR2_", "WT", "_", "Snord116", ".csv"), row.names = FALSE)

#check if dR2 genes is a subset of rhyI/II
RhyI.II = as.character(DCP_rhythm$rhythm.joint$gname)[DCP_rhythm$rhythm.joint$TOJR %in% c("rhyI", "rhyII")]
g.dR2 = as.character(DCP_dR2$gname)[DCP_dR2$p.R2<0.05]
all(g.dR2%in%RhyI.II)
```

```{bash runMetaScape}
##start docker
sudo systemctl start docker
##test run docker
sudo docker run hello-world

cd /home
cd xix66/msbio_v3
#I don't know why but I cannot run without sudo
#1. open the MSBio containers
sudo bin/up.sh 
#2. run analysis
sudo bin/ms.sh -o /data/CircadianPipeline/Methylation/MetaScape_Output/ByGeneTOJR2_rerun_mouse -S 10090 -T 10090 /data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_TOJR2_WT_Snord116.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json
# I suppose 10090 is code for mouse? https://metascape.org/gp/index.html#/menu/msbio. 

#shut down the container
sudo bin/down.sh
```

#### Pathway genes
```{r}
load("DCP_output.rData")
MAPK.genes = readxl::read_xlsx("MetaScape_Output/ByGeneTOJR2_MAPKgenes.xlsx") %>% as.data.frame()
colnames(MAPK.genes)[1] = "genes"
DCP_rhythm$gname_overlap = toupper(DCP_rhythm$gname_overlap)
DCP_rhythm$x1$gname = toupper(DCP_rhythm$x1$gname)
DCP_rhythm$x2$gname = toupper(DCP_rhythm$x2$gname)
DCP_rhythm$x1$rhythm$gname = toupper(DCP_rhythm$x1$rhythm$gname)
DCP_rhythm$x2$rhythm$gname = toupper(DCP_rhythm$x2$rhythm$gname)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = MAPK.genes %>% 
                                     filter(WT==1&Snord116!=1&genes%in%DCP_rhythm$gname_overlap) %>%pull(genes), 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/MAPKupGenes", file.height = 8, file.width = 16)

DiffCircaPipeline::DCP_PlotDisplay(DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Ereg", Info1 = "WT", Info2 = "Snord116"))
DiffCircaPipeline::DCP_PlotDisplay(DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Wdr77", Info1 = "WT", Info2 = "Snord116"))


```

#### Other outputs
```{r}
a.TOJR = DCP_rhythm$rhythm.joint$TOJR
tab.p = data.frame(cutoff = paste0("p-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy"))
a.TOJR = DCP_rhythm$rhythm.joint$TOJR.FDR
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.1, adjustP = FALSE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("p-value < ", 0.1),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.2, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.2),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.3, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.3),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
tab.p2 = tab.p[c(1, 3, 4), ]
colnames(tab.p2)[2] = "RhyI \n(WT)"
colnames(tab.p2)[3] = "RhyII \n(Snord116+/-)"
pdf(paste0(hdir, "/ByGene_TOJRtab.pdf"), width = 5)
p<-gridExtra::tableGrob(tab.p2, rows = NULL)
gridExtra::grid.arrange(p)
dev.off()


```

<!-- #### PSEA -->
<!-- The PSEA app is here: /Users/xiangningxue/Library/CloudStorage/OneDrive-UniversityofPittsburgh/Research/CommonMind/Control_Rhythmicity_Compare_Gender_rm2/PSEA  -->

<!-- I downloaded the mouse gene set here: http://ge-lab.org/#/data -->
<!-- ```{r} -->
<!-- PSEA.dir = paste0(hdir, "/PSEA/input") -->
<!-- dir.create(PSEA.dir, recursive = TRUE) -->
<!-- peak1 = DCP_rhythm$x1$rhythm %>%  -->
<!--   filter(gname %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR%in%c("rhyI", "both")) %>% pull("gname") %>% as.character)) %>% -->
<!--   select(c("gname", "peak")) -->
<!-- peak2 = DCP_rhythm$x2$rhythm %>%  -->
<!--   filter(gname %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR%in%c("rhyII", "both")) %>% pull("gname") %>% as.character)) %>% -->
<!--   select(c("gname", "peak")) -->
<!-- peak1$gname = toupper(peak1$gname) -->
<!-- peak2$gname = toupper(peak2$gname) -->
<!-- write.table(peak1, file = paste0(PSEA.dir, "/", "ByGene_WT", "_RhythGenes_Peak.txt"), row.names=FALSE, sep="\t", quote = FALSE, col.names = FALSE) -->
<!-- write.table(peak2, file = paste0(PSEA.dir, "/", "ByGene_Snord116", "_RhythGenes_Peak.txt"), row.names=FALSE, sep="\t", quote = FALSE, col.names = FALSE) -->

<!-- ``` -->

```{r, eval=FALSE}
m.res = mget(load("DCP_output.rData"))
```

#### Pathway genes corrected
```{r}
load("DCP_output.rData")
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Wdr77", 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/Wdr77.pdf", file.height = 4, file.width = 8)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Ube2q1", 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/Ube2q1.pdf", file.height = 4, file.width = 8)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Cited2", 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/Cited2.pdf", file.height = 4, file.width = 8)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Pdia3", 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/Pdia3.pdf", file.height = 4, file.width = 8)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = "Utp14b", 
                                   Info1 = "WT", Info2 = "Snord116",
                                   filename = "Figure/Utp14b.pdf", file.height = 3, file.width = 6)

```

### Weak signal discussion
```{r}
load("DCP_output.rData")
background.genes = x1$gname
length.max = length(background.genes)
fill.vec = function(a.vec, length.max=nrow(res)){
  c(a.vec, rep("", length.max-length(a.vec)))
}
df = data.frame(R2up = fill.vec(as.character(DCP_dR2$gname)[DCP_dR2$q.R2<0.4&DCP_dR2$delta.R2>0], length.max),
                R2down = fill.vec(as.character(DCP_dR2$gname)[DCP_dR2$q.R2<0.4&DCP_dR2$delta.R2<0], length.max),
                 Background = background.genes)
colnames(df) = c("R2up", "R2down", "_BACKGROUND")
write.csv(df, paste0("MetaScape_Input/DCP_byGene_TOJR0.5_dR2_q0.4", ".csv"), row.names = FALSE)
```


```{bash runMetaScape}
##start docker
sudo systemctl start docker
##test run docker
sudo docker run hello-world

cd /home
cd xix66/msbio_v3
#I don't know why but I cannot run without sudo
#1. open the MSBio containers
sudo bin/up.sh 
#2. run analysis
# for one analyses
<!-- sudo bin/ms.sh -o /data/CircadianPipeline/Methylation/MetaScape_Output/ByGeneTOJR /data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_TOJR_WT_Snord116.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json -->
sudo bin/ms.sh -o /data/CircadianPipeline/Methylation/MetaScape_Output/ByGeneTOJR2 /data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_TOJR2_WT_Snord116.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json
sudo bin/ms.sh -o data/CircadianPipeline/Methylation/MetaScape_Output/ByGenedR2 data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_dR2_WT_Snord116_2.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json
#re-run (previous version has a as.character problem)
sudo bin/ms.sh -o /data/CircadianPipeline/Methylation/MetaScape_Output/ByGeneTOJR2_rerun /data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_TOJR2_WT_Snord116.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json
#re-run 2 (previous forgot to specify this should be mouse )
sudo bin/ms.sh -o /data/CircadianPipeline/Methylation/MetaScape_Output/ByGeneTOJR2_rerun_mouse -S 10090 -T 10090 /data/CircadianPipeline/Methylation/MetaScape_Input/DCP_byGene_TOJR2_WT_Snord116.csv --option /data/CircadianPipeline/Methylation/MetaScape_Input/option.json
# I suppose 10090 is code for mouse? https://metascape.org/gp/index.html#/menu/msbio. 

# for multiple analyses #but only the first one is run
# sudo bin/ms.sh /data/CircadianPipeline/KyleRes/MetaScape_Input/CPonly.job
#shut down the container
sudo bin/down.sh
```


```{r}
hist(DCP_dR2$p.R2, breaks = seq(0, 1, 0.05))
p = ggplot(data.frame(x = DCP_dR2$p.R2), aes(x = x))+
  geom_histogram(color = "white", breaks=seq(0, 1, 0.05))+
  theme_bw()+
  xlab("p-value")+
  ggtitle("Histogram of p-values of the DRF test")+
  theme(axis.text.x = element_text(size = 14))

pdf("Figure/TOJR_0.05_dR2_pval_histogram.pdf", height = 4, width = 4)
print(p)
dev.off()
```

```{r}
m.res = mget(load("DCP_output.rData"))
DRtable = DiffCircaPipeline::SummarizeDR(result = list(m.res$DCP_dparam2$pvalue, m.res$DCP_dR2$p.R2, m.res$DCP_dR2$p.R2, m.res$DCP_dR2$p.R2, m.res$DCP_dR2$q.R2),
            test = c("phase", rep("DRF", 3), "DRF"),
            type = c(rep("p-value", 4), "q-value"),
            val = c(0.05, 0.1, 0.05, 0.01, 0.3))
DRtable = DRtable[, c(2, 3, 1)]
pdf(paste0("Figure/ByGene_DRresult.pdf"), width = 5)
p<-gridExtra::tableGrob(DRtable, rows = NULL)
gridExtra::grid.arrange(p)
dev.off()
```

