---
title: "SCZvsControl"
author: "Xiangning Xue"
date: '2022-06-08'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## prepare data
```{r}
meta = read.csv(paste0("data/CM_clinical_150_0514.csv"), row.names = 1)
dat = read.csv(paste0("data/CM_filtered_log2CPM_Combat_150_0514.csv"), row.names = 1)

meta.use = meta %>% filter(Dx=="SCZ"|matchIndex==1)
table(meta.use$Dx, meta.use$Gender)
dat.use = dat[, meta.use$DLPFC_RNA_Sequencing_Sample_ID]

library(biomaRt)
mart = useMart("ensembl")
listDatasets(mart) %>% filter(grepl("hsapien", dataset)) #hsapiens_gene_ensembl
mart = useDataset("hsapiens_gene_ensembl", mart)
listAttributes(mart)%>% filter(grepl("symbol", name)) #hgnc_symbol
listFilters(mart) %>% filter(grepl("ensembl", name)) #ensembl_gene_id
gNameMatch = getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), 
                   filters = "ensembl_gene_id", 
                   values = rownames(dat.use), 
                   mart)
dim(gNameMatch)
dim(dat.use)
gNameMatch2 = gNameMatch %>% right_join(data.frame(ensembl_gene_id = rownames(dat.use), ensembl_gene_id_origin = rownames(dat.use)))
gNameMatch2 = gNameMatch2[!duplicated(gNameMatch2$ensembl_gene_id), ]
gNameMatch2$hgnc_symbol[is.na(gNameMatch2$hgnc_symbol)|gNameMatch2$hgnc_symbol==""] = gNameMatch2$ensembl_gene_id[is.na(gNameMatch2$hgnc_symbol)|gNameMatch2$hgnc_symbol==""]
```

## DCP 
```{r}
x1 = list(data = data.frame(dat.use[, meta.use$Dx =="Control"]), 
          time = meta.use[meta.use$Dx =="Control", "Corrected.TOD"], 
          gname = gNameMatch2$hgnc_symbol[match(rownames(dat.use), gNameMatch2$ensembl_gene_id_origin)])
x2 = list(data = data.frame(dat.use[, meta.use$Dx =="SCZ"]), 
          time = meta.use[meta.use$Dx =="SCZ", "Corrected.TOD"], 
          gname = gNameMatch2$hgnc_symbol[match(rownames(dat.use), gNameMatch2$ensembl_gene_id_origin)])

DCP_rhythm = DiffCircaPipeline::DCP_Rhythmicity(x1, x2, alpha = 0.05, alpha.FDR = 0.05)

table(DCP_rhythm$rhythm.joint$TOJR)
table(DCP_rhythm$rhythm.joint$TOJR.FDR)
DCP_dparam = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "A&phase")
DCP_dparam2 = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "phase")
DCP_dparam3 = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "A")
DCP_dR2 = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm)
DCP_dR2_perm = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm, method = "permutation", nSampling = 1000, Sampling.save = "perm0.05", parallel.ncores = 20)
plot(-log10(DCP_dR2$p.R2), -log10(DCP_dR2_perm$p.R2))
abline(a = 0, b = 1, col = "red")
# save(x1, x2, DCP_rhythm, DCP_dparam, DCP_dparam2, DCP_dR2, DCP_dR2_perm, file = paste0("DCP_output.rData"))
```


### outputs
```{r, echo=FALSE}
load("DCP_output.rData")

a.TOJR = DCP_rhythm$rhythm.joint$TOJR
tab.p = data.frame(cutoff = paste0("p-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy"))
a.TOJR = DCP_rhythm$rhythm.joint$TOJR.FDR
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.1, adjustP = FALSE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("p-value < ", 0.1),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.2, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.2),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.3, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.3),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
tab.p2 = tab.p[c(1, 3, 4), ]
colnames(tab.p2)[2] = "RhyI \n(Control)"
colnames(tab.p2)[3] = "RhyII \n(SCZ)"

pdf(paste0("Figures/TOJRtab.pdf"), width = 5)
p<-gridExtra::tableGrob(tab.p2, rows = NULL)
gridExtra::grid.arrange(p)
dev.off()

DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dR2 %>% arrange(p.R2) %>%
                                     filter(p.R2<0.05&delta.R2>0) %>% pull(gname),
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_dR2up.pdf"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dR2 %>% arrange(p.R2) %>%
                                     filter(p.R2<0.05&delta.R2<0) %>% pull(gname),
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_dR2down.pdf"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_PlotPeakDiff(DCP_rhythm, dPhase = DCP_dparam2)


```


### IPA
```{r}
df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.05&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.05&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/dR2.csv"), row.names = FALSE)

df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.05&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.05&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/dR2_perm.csv"), row.names = FALSE)

df = data.frame(ID = background.genes,
                Control = as.numeric(!background.genes %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR %in% c("rhyI", "both")) %>% pull(gname))),
                SCZ = as.numeric(!background.genes %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR %in% c("rhyII", "both")) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR0.05_TOJR.csv"), row.names = FALSE)

#dR2 p<0.1
df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.1&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.1&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/dR2_p0.1.csv"), row.names = FALSE)
```

### Pathway genes
```{r}
#LPS/IL-1 Mediated Inhibition of RXR Function
LPS.genes = c("ACOX3","CAT","CYP2J2")
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = LPS.genes,
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_TOJR0.05_dR20.1_down_LPS.pdf"), file.height = 4, file.width = 8)
Calcium.genes = c("CACNA1F", "HDAC8", "MYL6B", "NFATC4")
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = Calcium.genes,
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_TOJR0.05_dR20.1_up_calcium.pdf"), file.height = 4, file.width = 8)

```


## DCP with cutoff 0.1
```{r}
load("DCP_output.rData")
DCP_rhythm = DiffCircaPipeline::DCP_Rhythmicity(x1, x2, alpha = 0.1, alpha.FDR = 0.1)

table(DCP_rhythm$rhythm.joint$TOJR)
table(DCP_rhythm$rhythm.joint$TOJR.FDR)
DCP_dparam = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "A&phase")
DCP_dparam2 = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "phase")
DCP_dparam3 = DiffCircaPipeline::DCP_DiffPar(DCP_rhythm, Par = "M")
DCP_dR2 = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm)
DCP_dR2_perm = DiffCircaPipeline::DCP_DiffR2(DCP_rhythm, method = "permutation", nSampling = 10000, Sampling.save = "perm0.1", parallel.ncores = 20)
#save(x1, x2, DCP_rhythm, DCP_dparam, DCP_dparam2, DCP_dparam3, DCP_dR2,DCP_dR2_perm, file = paste0("DCP_output_0.1.rData"))
```


### outputs
```{r, echo=FALSE}
load("DCP_output_0.1.rData")

a.TOJR = DCP_rhythm$rhythm.joint$TOJR
tab.p = data.frame(cutoff = paste0("p-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy"))
a.TOJR = DCP_rhythm$rhythm.joint$TOJR.FDR
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.05),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.1, adjustP = FALSE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("p-value < ", 0.1),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.2, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.2),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
a.TOJR = DiffCircaPipeline::toTOJR(DCP_rhythm, method = "Sidak_FS", alpha = 0.3, adjustP = TRUE, parallel.ncores = 2)
tab.p = rbind.data.frame(tab.p, 
                         data.frame(cutoff = paste0("q-value < ", 0.3),
                   RhyI = sum(a.TOJR=="rhyI"),
                   RhyII = sum(a.TOJR=="rhyII"),
                   RhyBoth = sum(a.TOJR=="both"),
                   Arrhy = sum(a.TOJR=="arrhy")))
tab.p2 = tab.p[c(1, 3, 4), ]
colnames(tab.p2)[2] = "RhyI \n(Control)"
colnames(tab.p2)[3] = "RhyII \n(SCZ)"

pdf(paste0("Figures/TOJRtab.pdf"), width = 5)
p<-gridExtra::tableGrob(tab.p2, rows = NULL)
gridExtra::grid.arrange(p)
dev.off()

DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dR2 %>% arrange(p.R2) %>%
                                     filter(p.R2<0.05&delta.R2>0) %>% pull(gname),
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_dR2up.pdf"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = DCP_dR2 %>% arrange(p.R2) %>%
                                     filter(p.R2<0.05&delta.R2<0) %>% pull(gname),
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_dR2down.pdf"), file.height = 8, file.width = 16)
DiffCircaPipeline::DCP_PlotPeakDiff(DCP_rhythm, dPhase = DCP_dparam2)


```



### Summarize DR results 
```{r, eval=FALSE}
SC.res = mget(load("DCP_output.rData"))
SummarizeDR = function(result = DCP_dR2$p.R2, test = "DRF", type = "p-value", val = 0.05, out = "long"){
  if(class(result)!="list"){
    result = list(result)
  }
  l = length(result)
  stopifnot("Arguments `result`, `test`, `type`, `val` must have the same length" = (length(test)==l&length(type)==l&length(val)==l))
  tab = lapply(1:l, function(a){
    a.result = result[[a]]
    a.test = test[a]
    a.type = type[a]
    a.val = val[a]
    a.tab0 = table(a.result<a.val)
    a.tab = data.frame(nTRUE = unname(ifelse(is.na(a.tab0["TRUE"]), 0, a.tab0["TRUE"])),
                       nFALSE = unname(ifelse(is.na(a.tab0["FALSE"]), 0, a.tab0["FALSE"])),
                       test = a.test,
                       cutoff = paste0(a.type, "<", a.val))
    return(a.tab)
  })
  tab = do.call(rbind.data.frame, tab)
  tab$nTRUE = paste0(tab$nTRUE, "(/", tab$nTRUE+tab$nFALSE, ")")
  if(out=="long"){
    return(tab[, c(1, 3,4)])
  }else if(out =="wide"){
    tab.list = lapply(1:nrow(tab[, c(1, 3,4)]), function(i){
      a.cell = data.frame(x = c(tab[i, "nTRUE"]))
      rownames(a.cell) = NULL
      colnames(a.cell) = paste0(tab[i, "test"], " ", tab[i, "cutoff"])
      return(a.cell)
    })
    tab2 = do.call(cbind.data.frame, 
                   tab.list)
    return(tab2)
  }
}

#SCZ
DRtable = DiffCircaPipeline::SummarizeDR(result = list(SC.res$DCP_dR2$p.R2, SC.res$DCP_dR2$p.R2, SC.res$DCP_dR2$q.R2),
            test = c(rep("DRF", 2), "DRF"),
            type = c(rep("p-value", 2), "q-value"),
            val = c(0.05, 0.1, 0.4), 
            "long")
pdf(paste0("Figures/DRFresultSummary.pdf"), width = 5)
p<-gridExtra::tableGrob(DRtable, rows = NULL)
gridExtra::grid.arrange(p)
dev.off()
```

### IPA
```{r}
df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.05&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2 %>% filter(p.R2<0.05&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR0.1_dR2.csv"), row.names = FALSE)

df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.05&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.05&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR0.1_dR2_perm.csv"), row.names = FALSE)

df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.1&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2_perm %>% filter(p.R2<0.1&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR0.1_dR2_perm0.1.csv"), row.names = FALSE)

df = data.frame(ID = background.genes,
                Control = as.numeric(background.genes %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR %in% c("rhyI", "both")) %>% pull(gname))),
                SCZ = as.numeric(background.genes %in% (DCP_rhythm$rhythm.joint %>% filter(TOJR %in% c("rhyII", "both")) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR0.1_TOJR.csv"), row.names = FALSE)

df = data.frame(ID = CMres$genes,
                up = as.numeric(CMres$genes %in% (CMres2 %>% filter(R2gainPvalue<0.05) %>% pull(genes))),
                down = as.numeric(CMres$genes %in% (CMres2 %>% filter(R2losePvalue<0.05) %>% pull(genes)))
                )
write.csv(df, paste0("IPA_Input/NatureComm_dR2.csv"), row.names = FALSE)


```

#### pathways
```{r}
#Mitochondrial dysfunction 
mito.genes = "ATP5F1A,ATP5MC1,ATP5MC2,ATP5MG,ATPAF1,COX6A1,COX7A1,COX8A,HTRA2,NDUFB11,NDUFB2,NDUFS2,NDUFS3,NDUFS5,NDUFS8,NDUFV1,PARK7,PRDX5,PSEN2,UQCRC1,UQCRFS1"
mito.genes = unlist(strsplit(mito.genes, ","))
DCP_rhythm$rhythm.joint[match(mito.genes, DCP_rhythm$rhythm.joint$gname),]
DCP_dR2_perm[match(mito.genes, DCP_dR2_perm$gname),]

DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = mito.genes,
                                  Info1 = "Control", Info2 = "SCZ", 
                                  filename = "SCZvsControl/Pathway_MitochondroalDysfunction_DCPplot.pdf", file.width = 14)
plot(DCP_dR2$R2.2-DCP_dR2$R2.1, -log10(DCP_dR2$p.R2), col = factor(DCP_dR2$p.R2<0.2))
```

#### Make enrichment plots
```{r}
IPA_R2down = readxl::read_xls("IPA_Output/TOJR_0.05_dR2_0.1/CanonicalPathways_R2down.xls", 
                              skip = 1)
IPA_R2up = readxl::read_xls("IPA_Output/TOJR_0.05_dR2_0.1/CanonicalPathways_R2up.xls", 
                              skip = 1)

# read.table("IPA_Output/TOJR0.05_dR2_0.1/CanonicalPathways_R2down.xls", fill = TRUE, skip = 2, header = TRUE, sep = "\t")[, 1:3]
# IPA_SCZ = read.table("IPA_Output/TOJR0.1_TOJR/CanonicalPathways_SCZ.txt", fill = TRUE, skip = 2, header = TRUE, sep = "\t")[, 1:3]

IPA_R2down$nOverlap = sapply(strsplit(IPA_R2down$Molecules, ","), length)
IPA_R2up$nOverlap = sapply(strsplit(IPA_R2up$Molecules, ","), length)
IPA_R2down2 = IPA_R2down %>% filter(`-log(p-value)`>-log10(0.05)&nOverlap>=3)
IPA_R2up2 = IPA_R2up %>% filter(`-log(p-value)`>-log10(0.05)&nOverlap>=3)
library(ggplot2)
length(intersect(IPA_R2down2$`Ingenuity Canonical Pathways`, IPA_R2up2$`Ingenuity Canonical Pathways`))#no overlap
# df = data.frame(Pathways = IPA_Control2$Ingenuity.Canonical.Pathways, UC = IPA_Control2$X.log.p.value.) %>% full_join(data.frame(Pathways = IPA_SCZ2$Ingenuity.Canonical.Pathways, SCZ = IPA_SCZ2$X.log.p.value.)) %>% 
#   mutate(UC = ifelse(is.na(UC), 0, UC), 
#          SCZ = ifelse(is.na(SCZ), 0, SCZ)) %>% 
#   tidyr::pivot_longer(2:3, names_to = "Group", values_to = "-log10(p-value)") %>%
#   mutate(Group = factor(Group))

df = rbind.data.frame(IPA_R2down2 %>% mutate(Group = "decrease"), 
                      IPA_R2up2 %>% mutate(Group = "increase")) %>% 
  mutate(Group = factor(Group)) %>% 
  arrange(`-log(p-value)`)%>%
  arrange(Group)
df$order = rownames(df)


p = ggplot(df, aes(x = reorder(`Ingenuity Canonical Pathways`,order,
                           function(x)as.numeric(x)), y = `-log(p-value)`, fill = Group))+
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept=-log10(0.05), color = "#ede6e6")+
  scale_fill_manual(values = c("#0da105", "#0512a1"), 
                    breaks = c("decrease", "increase"), 
                    labels = c(expression(Delta*R^2~"<"*0), 
                               expression(Delta*R^2~">"*0)))+
  xlab("")+
  ylab(expression(-log[10]*"(p-value)"))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(face="bold")) + 
  coord_flip()
pdf("Figures/dR2_IPA_plot.pdf", height = 5)
print(p)
dev.off()
```

## Weak signal discussion

### IPA with qval<0.4
```{r}
load("DCP_output.rData")
background.genes = x1$gname
df = data.frame(ID = background.genes,
                R2up = as.numeric(background.genes %in% (DCP_dR2 %>% filter(q.R2<0.4&delta.R2>0) %>% pull(gname))),
                R2down = as.numeric(background.genes %in% (DCP_dR2 %>% filter(q.R2<0.4&delta.R2<0) %>% pull(gname)))
                )
write.csv(df, paste0("IPA_Input/TOJR_p0.5_dR2_q0.4.csv"), row.names = FALSE)

```



```{r}
IPA_R2down = readxl::read_xls("IPA_Output/TOJR_p0.05_dR2_q0.4/CanonicalPathways_R2down.xls", 
                              skip = 1)
IPA_R2up = readxl::read_xls("IPA_Output/TOJR_p0.05_dR2_q0.4/CanonicalPathways_R2up.xls", 
                              skip = 1)

# read.table("IPA_Output/TOJR0.05_dR2_0.1/CanonicalPathways_R2down.xls", fill = TRUE, skip = 2, header = TRUE, sep = "\t")[, 1:3]
# IPA_SCZ = read.table("IPA_Output/TOJR0.1_TOJR/CanonicalPathways_SCZ.txt", fill = TRUE, skip = 2, header = TRUE, sep = "\t")[, 1:3]

IPA_R2down$nOverlap = sapply(strsplit(IPA_R2down$Molecules, ","), length)
IPA_R2up$nOverlap = sapply(strsplit(IPA_R2up$Molecules, ","), length)
IPA_R2down2 = IPA_R2down %>% filter(`-log(p-value)`>-log10(0.05)&nOverlap>=3)
IPA_R2up2 = IPA_R2up %>% filter(`-log(p-value)`>-log10(0.05)&nOverlap>=3)

IPA_R2down %>% mutate(noverlap = length(strsplit(Molecules, ",")))
library(ggplot2)
length(intersect(IPA_R2down2$`Ingenuity Canonical Pathways`, IPA_R2up2$`Ingenuity Canonical Pathways`))#no overlap
# df = data.frame(Pathways = IPA_Control2$Ingenuity.Canonical.Pathways, UC = IPA_Control2$X.log.p.value.) %>% full_join(data.frame(Pathways = IPA_SCZ2$Ingenuity.Canonical.Pathways, SCZ = IPA_SCZ2$X.log.p.value.)) %>% 
#   mutate(UC = ifelse(is.na(UC), 0, UC), 
#          SCZ = ifelse(is.na(SCZ), 0, SCZ)) %>% 
#   tidyr::pivot_longer(2:3, names_to = "Group", values_to = "-log10(p-value)") %>%
#   mutate(Group = factor(Group))

df = rbind.data.frame(IPA_R2down2 %>% mutate(Group = "decrease"), 
                      IPA_R2up2 %>% mutate(Group = "increase")) %>% 
  mutate(Group = factor(Group)) %>% 
  arrange(`-log(p-value)`)%>%
  arrange(Group)
df$order = rownames(df)


p = ggplot(df, aes(x = reorder(`Ingenuity Canonical Pathways`,order,
                           function(x)as.numeric(x)), y = `-log(p-value)`, fill = Group))+
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept=-log10(0.05), color = "#ede6e6")+
  scale_fill_manual(values = c("#0da105", "#0512a1"), 
                    breaks = c("decrease", "increase"), 
                    labels = c(expression(Delta*R^2~"<"*0), 
                               expression(Delta*R^2~">"*0)))+
  xlab("")+
  ylab(expression(-log[10]*"(p-value)"))+
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.line = element_line(colour = "black"),
        axis.text.y = element_text(face="bold")) + 
  coord_flip()
pdf("Figures/TOJR_p0.5_dR2_q0.4_IPA_plot.pdf", height = 5)
print(p)
dev.off()

# 
# 
# table(DCP_dR2$q.R2<0.4, DCP_dR2$delta.R2>0)
# hist(DCP_dR2$p.R2, breaks = seq(0, 1, 0.05))

```

```{r}
hist(DCP_dR2$p.R2, breaks = seq(0, 1, 0.05))
p = ggplot(data.frame(x = DCP_dR2$p.R2), aes(x = x))+
  geom_histogram(color = "white", breaks=seq(0, 1, 0.05))+
  theme_bw()+
  xlab("p-value")+
  ggtitle("Histogram of p-values of the DRF test")
pdf("Figures/TOJR_0.05_dR2_pval_histogram.pdf", height = 5)
print(p)
dev.off()
```

### Pathway genes
```{r}
#LPS/IL-1 Mediated Inhibition of RXR Function
Pulmonary.genes = c("CXCL12","ETV5","HBEGF","TCF7")
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = Pulmonary.genes,
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_TOJR0.05_dR2q0.4_down_Pulmonary.pdf"), file.height = 4, file.width = 8)
Cholesterol.genes = c("EBP","HSD17B7","MVD")
DiffCircaPipeline::DCP_ScatterPlot(DCP_rhythm, genes.plot = Cholesterol.genes,
                                   Info1 = "Control", Info2 = "SCZ",
                                   filename = paste0("Figures/scatter_TOJR0.05_dR2q0.4_up_Cholesterol.pdf"), file.height = 4, file.width = 8)

```

